Extracting The Power of Dependent Types

### 2021-09-21 paper.tex
Restore 12 pages limit
Trivial fixes.

\documentclass[sigplan,screen,anonymous,review]{acmart}

@@

  Our main insight is that it is possible to
  express all three steps in a single language that supports
  both dependent typ\-es and reflection. Essentially,
  this allows us to express a program, its formal properties, and
  a compiler for it hand-in-hand, offering a lot of flexibility to programmers.
  We demonstrate this three-step approach by embedding a subset of
  the PostScript language in Agda, and illustrating it with several short examples.
  Thus we use the power of reflection to bring the benefits of dependent types to languages
  that had to go without them so far.

%  Dependently-typed host languages empower users to verify a wide
%  range of properties of embedded languages and
%  programs written in them.  Designers of such embedded languages
%  are faced with a difficult choice between using a shallow or a deep
%  embedding.  The former is easier to use because the entire
%  infrastructure of the host langauge is immediately
%  available. Meanwhile, the latter gives full access to the
%  structure of embedded programs, but is difficult to use in practice,
%  especially when the embedded language is itself dependently typed.
%
%  The main insight presented in this paper is that the choice between
%  shallow and deep embedding can be eliminated by working in a host
%  language with reflection capabilities: we start from a shallow
%  embedding that can use all libraries and tools of the host language,
%  and later use reflection to expose the deep structure of the
%  embedded programs.
%  %
%  Concretely, we apply this technique to embed three
%  programming languages --- Kaleidoscope, SaC, and (a subset of) APL ---
%  into the dependently typed theorem prover Agda, using dependent
%  types to statically enforce several properties of interest.  We then
%  use Agda's reflection capabilities to extract the embedded programs
%  back into the original language, so that the existing toolchain can
%  be leveraged.
%  %
%  In this process, statically verified properties of the host language
%  are mapped onto runtime checks in the target language, allowing
%  extracted programs to interact safely with existing code.
%  %
%  Finally, we demonstrate the feasibility of our approach with the
%  implementation and extraction of a convolutional neural network in
%  our embedding of APL.\@

%  Most practical programming languages of today prioriterise
%  performance and usability over safety.  As a result, many
%  useful correctness invariants cannot be expressed in the given
%  language, typically due to weakness of the chosen type system.
%  We are looking for non-invasive verification: how to
%  guarantee a property about the given program without any modifications
%  to the language toolchain.
%
%  We study an approach when a programming language is shallowly
%  embedded into a theorem prover, and dependent types are used
%  to encode the properties of interest.   Bringing the embedded
%  langauge and the properties back into the original language
%  is the main focus of this work.  We propose a generic extraction
%  mechanism that we implement in Agda that is based on the concept
%  of reflection.  Making it possible to translate any shallowly
%  embedded language in the target backend without the necessity
%  to modify the theorem prover.
%  We demonstrate our framework in action by encoding three languages:
%  Kaleidoskope, SaC and a subset of APL, culminating in the implementation
%  and extraction of a convolutional neural network.
\end{abstract}
\maketitle

@@

%\paragraph{Metaprogramming}
There is a large body of
work on metaprogramming facilities in various programming
languages.  \citet{refl-masses} track the origins
of metaprogramming to Smith's work on reflection in Lisp \cite{refl-lisp}.
Some prominent metaprogramming systems include
MetaOcaml \cite{metaocaml}, Me\-taML \cite{metaml},
reFlect \cite{DBLP:journals/jfp/GrundyMO06},
Template Haskell \cite{sheard2002template},
Racket \cite{plt-tr1}, and various other Lisp/Scheme dialects.
%
However, these systems typically do not support dependent
types, so they are not well suited for our goal of statically
enforcing correctness of embedded programs.

@@

%\paragraph{Extraction}
The Coq proof assistant is equipped with extraction
capabilities~\cite{10.1007/978-3-540-69407-6_39,10.1007/3-540-39185-1_12},
which extracts functional code from Coq proofs (or programs).  The
default target language is Ocaml, but a few other options were added
recently.
%
Likewise, Agda itself has a mechanism for defining custom backends, of
which the GHC backend is the most prominent.
%
Other proof assistants provide similar extraction tools as well.
%
The main difference from our approach in this paper
is that these extractors are written as plugins to the proof
assistant, while we implement our extractors directly in the proof
assistant itself.
%
While it would be possible to implement extractors presented in this
paper as Coq plugins or Agda backends, conceptually they are more
heavyweight.  Our extractors and programs can (in principle)
communicate with each other. In addition, as they are just Agda programs, they can
be reflected themselves and their structure can be leveraged.

@@

%% Bibliography
\bibliography{paper}

>>>>

\documentclass[sigplan,screen]{acmart}

@@

  Our main insight is that it is possible to
  express all three steps in a single language that supports
  both dependent typ\-es and reflection. Essentially,
  this allows us to express a program, its formal properties, and
  a compiler for it hand-in-hand, offering a lot of flexibility to programmers.
  We demonstrate this three-step approach by embedding a subset of
  the PostScript language in Agda, and illustrating it with several short examples.
  Thus we use the power of reflection to bring the benefits of dependent types to languages
  that had to go without them so far.
\end{abstract}
\maketitle

@@

%\paragraph{Metaprogramming}
There is a large body of
work on metaprogramming facilities in various programming
languages.  \citet{refl-masses} track the origins
of metaprogramming to Smith's work on reflection in Lisp \cite{refl-lisp}.
Some prominent metaprogramming systems include
MetaOCaml \cite{metaocaml}, Me\-taML \cite{metaml},
reFlect \cite{DBLP:journals/jfp/GrundyMO06},
Template Haskell \cite{sheard2002template},
Racket \cite{plt-tr1}, and various other Lisp/Scheme dialects.
%
However, these systems typically do not support dependent
types, so they are not well suited for our goal of statically
enforcing correctness of embedded programs.

@@

%\paragraph{Extraction}
The Coq proof assistant is equipped with extraction
capabilities~\cite{10.1007/978-3-540-69407-6_39,10.1007/3-540-39185-1_12},
which extracts functional code from Coq proofs (or programs).  The
default target language is OCaml, but a few other options were added
recently.
%
Likewise, Agda itself has a mechanism for defining custom backends, of
which the GHC backend is the most prominent.
%
Other proof assistants provide similar extraction tools as well.
%
The main difference from our approach in this paper
is that these extractors are written as plugins to the proof
assistant, while we implement our extractors directly in the proof
assistant itself.
%
While it would be possible to implement extractors presented in this
paper as Coq plugins or Agda backends, conceptually they are more
heavyweight.  Our extractors and programs can (in principle)
communicate with each other. In addition, as they are just Agda programs, they can
be reflected themselves and their structure can be leveraged.

@@

\pagebreak
%% Bibliography
\bibliography{paper}
